Demonstrating the Reliability of Randomized Measurement and Verification for Switchable Control Retrofits Using a Large Open-Source Dataset

### 2025-03-22 code/manuscript.Rmd
Uploaded revised manuscript, added sampling interval timeline

---
title: "Demonstrating the Reliability of Randomized Measurement and Verification for Switchable Control Retrofits Using a Large Open-Source Dataset"
author:
  - Aoyu Zou ^[Center for the Built Environment, University of California Berkeley, USA], Paul Raftery ^1^, Stefano Schiavon ^1^, Carlos Duarte ^1^, Gail Brager ^1^
abstract: Conventional measurement and verification (M&V) methods rely on pre- and post- retrofit comparison to estimate energy savings. They are often time-consuming and unreliable, especially when non-routine events such as reduced occupancy, or equipment failures occur during the M&V process. Those events are unrelated to the intervention but significantly affect building energy consumption and thus when the analyst applies conventional M&V suggested by industry guidelines, the results can be largely confounded. In this study, we demonstrated that switchable interventions, such as most of the control retrofits, can benefit from a new M&V method which randomly samples whether to implement the baseline or the intervention strategy each day. We tested this novel randomized M&V method on a large public dataset covering various climate zones and commercial building types, using a virtual chilled water supply temperature reset based on outdoor weather as the intervention. Our results show that compared to the conventional method, the randomized method provides faster and more accurate savings estimation. Additionally we found that when non-routine events are present, the randomized M&V approach estimates savings that are much closer to the ground-truth savings than the conventional M&V method, demonstrating much improved reliability.

@@

# colors
ls_colors <- c("Baseline" = "#c6dbef",
               "Measured baseline" = "#c6dbef",
               "Adjusted baseline" = "#c6dbef",
               "Projected baseline" = "#2171b5",
               "Intervention" = "#fdbb84",
               "Measured interv" = "#fdbb84",
               "Conventional" = "grey70", 
               "Randomized" = "#99d8c9", 
               "Randomized\n(24 months)" = "#66c2a4",
               "Randomized\n(50/50)" = "#66c2a4",
               "Daily\nsampling" = "#66c2a4",
               "2-day\nsampling" = "#66c2a4",
               "3-day\nsampling" = "#66c2a4",
               "7-day\nsampling" = "#66c2a4",
               "Dropped" = "#41ae76",
               "Kept" = "#006d2c",
               "Randomized\n(20/80)" = "#ccece6",
               "Buildings finishing randomized M&V" =  "#99d8c9",
               "Buildings satisfying 80% TMY range" = "black", 
               "Buildings satisfying SPRT" = "black")

@@

Measurement and Verification (M&V) is the process of quantifying energy
savings from energy efficiency projects by comparing actual energy
consumption against a baseline, adjusting for factors like weather and
occupancy. This process ensures that improvements in energy performance
are accurately evaluated. In the United States, practitioners often
refer to ASHRAE Guideline 14, the International Performance Measurement
and Verification Protocol (IPMVP), and the Federal Energy Management
Program (FEMP) for standard guidelines [@doe_mv_2008;
@efficiency_valuation_organisation_international_2012;
@ashrae_ashrae_2023]. These guidelines outline standardized methods for
quantifying energy savings, whether through calibrated simulations or
monitored measurements for specific equipment or systems (isolation
methods) or for entire buildings (whole-building methods). In this
study, we will focus on the energy savings quantified at the
whole-building level where the measurements are obtained from utility
bills or whole-building meters, however the overall conclusions would
also apply to isolation methods. Typically, the process begins with
baseline measurements taken during the year before implementing the 
energy-efficiency retrofit, followed by the same measurement procedure
during the year after the intervention. After collecting two years of
data, an M&V analyst fits an energy prediction model, using variables
such as outdoor temperature and time [@mathieu_quantifying_2011] to
project baseline energy consumption in the post-retrofit period. The
difference between the counter-factual baseline and the measured
intervention period energy consumption represents the energy savings.
One of the key drawbacks of this method is its reliance on a two-year
timeline to quantify savings, during which baseline measurements can
become outdated due to changes in building performance caused by
non-routine events unrelated to the intervention. This limitation
reduces the feasibility of rapid M&V and complicates the quantification
of estimation uncertainty, thus impacting the accuracy and timeliness of
savings assessments.

@@

To address the limitations of conventional M&V methods and the
challenges posed by non-routine events, we propose a novel M&V method
that adopts the randomized crossover design, a gold standard from
medical and agricultural studies [@blackston2019; @duan2013; @gupta2019;
@mie2022; @munro; @raseduzzaman2017]. Another improvement is that we
proposed a sequential evaluation framework and defined stopping criteria
to end the M&V if the target effect is detected. This is to avoid
unnecessary measurement collection over the full 2-year M&V cycle. The
full framework is detailed in a previous published study with all
stopping criteria and example usecases outlined [@raftery2024]. In
summary, this method provides M&V analysts with a randomized schedule
that alternates between baseline and intervention implementation while
ensuring balanced sampling across days of the week and seasons. For
example, given a 10-week M&V period for 1 intervention, the balanced
randomized schedule would equally sample 5 Mondays with the baseline
strategy in operation and 5 Mondays with the intervention strategy in
operation. The limitation of this method is that it is only applicable
to a subset of retrofit projects where interventions can be easily
switched on and off. However, for all applicable use cases, it allows
analysts to detect energy savings sequentially right after test begins.
In addition, once the desired savings target is achieved, analysts can
terminate the M&V and switch to 100% intervention. The key advantage of
randomization is that if control strategies are sampled with equal
probability, the influence of other confounding factors such as
occupancy change (or commonly known as non-routine events) is likely to
be evenly distributed among measurements, leading to a more accurate and
unbiased assessment of the intervention's impact.

@@

The Building Genome Dataset 2 (BGD2) is an extensive open-access dataset
designed to advance research and development in building energy
efficiency and control strategies acting as a test-bed for modeling,
simulation, and algorithm development [@miller_building_2020]. BGD2
contains over 1000 buildings' metadata and realistic operational
measurements between 2016 and 2017 from across North America and Europe,
making it one of the most comprehensive collections of building-related
data available for scientific use. The dataset includes various
commercial building types such as offices, education facilities, public,
and retail buildings, and provides detailed information on their
physical characteristics (e.g. energy ratings, heating types and floor
area) and hourly measurements of chilled and hot water, electricity, gas
usage as well as site outdoor weather conditions. In this study, we
mostly use the metered electricity and outdoor weather measurements for
running the M&V methods.

@@

ASHRAE Guidelinge 14 and IPMVP provides options for whole building M&V
use cases. For code compliance using prescriptive option, 12 months of
baseline and 12 months of postretrofit measurements are required. In
addition, the expected savings should be larger than 10% and the
baseline model fitting accuracy should have CV(RMSE) lower than 20%.
Thus, most research related to M&V for whole-building approach focuses
on the accuracy of baseline modeling, exploring model performance from
simple regression models to more complex machine learning techniques.
One study reviewed various models suitable for M&V applications as well
as selected input features [@alrobaie_review_2022] and another study
provided a definitive methodology to apply machine learning models for
M&V use cases [@gallagher_development_2018]. In addition, a few studies
investigated the critical performance metrics to evaluate the developed
baseline models [@granderson_automated_2015;
@granderson_development_2014] and compared a variety of models using
those metrics [@granderson_accuracy_2016]. These studies made
significant contributions by emphasizing the uncertainty associated with
the model-fitting process, a key factor in accurately determining energy
savings. Furthermore, other researchers addressed this issue by
leveraging statistical formulation and inference to improve baseline
energy models [@burkhart_measurement_2014; @heo_calibration_2012;
@walter_uncertainty_2014]. However, a gap still remains in the
literature regarding the rigorous quantification of uncertainties
directly associated with calculated savings, for instance, accounting
for the potential bias that baseline model might deteriorate (i.e.
becomes 'stale') over an extended period of pre- and post-analysis.

@@

Non-routine events during M&V commonly refers to unexpected changes in a
building that influences its energy usage. These changes can greatly
affect measured energy consumption in buildings and are typically
unrelated to the intervention strategy. Therefore, those changes are
considered 'static factors' respect to measured independent variables. A
common non-routine event in energy-saving M&V projects is a change in
occupancy or a significant shift in occupant behavior, equipment run
time, and operating conditions (e.g. set points, lighting and
ventilation levels). However, current standards or guidelines only
provides general approach for consideration. For example, IPMVP requires
facility owner and the M&V analyst periodically perform inspections of
all equipment and operations during reporting period, which is labor
intensive and error-prone. ASHRAE guideline 14 recommends performing
engineering calculations or computer software simulations to adjust
post-retrofit baseline. Additional, it is very rare to have access to
all measurements such as occupancy, and thus the analyst normally assume
those factors remain unchanged throughout the study. Specifically,
studies investigating the effects of demand response on building energy
efficiency commonly uses linear interpolation to estimate
counter-factual baseline [@beil_round-trip_2015;
@keskar_commercial_2020], and one study points out that it is inaccurate
to assume no change in the operating conditions during the test period
[@huang_experimental_2023]. One likely encountered non-routine event due
to operation condition change is filter clogging in air handling units
due to particle accumulation. This can cause supply fans to gradually
consume more energy to maintain required duct static pressure
[@feng_newly_2019; @zhai_full-scale_2017]. If M&V analysts are unaware
of such changes and lack an appropriate adjustment method (e.g.,
replacing filters before the intervention begins), the savings could be
underestimated as increased energy use is incorrectly attributed to the
intervention rather than the mechanical issue. Some studies realized the
limitation of current M&V methods, which only consider adjusting for
outdoor weather, is insufficient and emphasized the importance of
requiring matched comparison groups to control for exogenous factors
beyond weather differences when comparing between baseline and
intervention [@demand_side_analytics_population_2022;
@huang_experimental_2023]. However, those methods are still unable to
accurately quantify uncertainties due to changing baseline.

@@

As mentioned, the goal of an M&V project is to determine the
effect---typically energy savings---of an switchable energy-efficient
intervention. The goal is to determine, for a large sample of buildings,
how much more accurately the novel M&V method would estimate the savings
of such an intervention compared to conventional M&V, and how much more
quickly it would reach a result. An example of such an intervention is a
control retrofit developed by a software-as-a-service company that
adjusts the chilled water plant's supply water temperature, which can be
commonly found in the literature [@duarte_field_2023;
@lee_simulationoptimization_2012; @qiu_chilled_2022; @jin2007;
@taylor_optimizing_2012]. In our case, we make it even simpler by
adjusting setpoint based on outdoor weather conditions with more
description shown in Section \@ref(apply-virtue-intervention).
Therefore, we defined the M&V scenario as follows:

"A company aims to sell its supply temperature reset control software
package to a customer, in this case, the building owner, with a
promotion that it will reduce the building's electricity usage. If the
building owner decides to purchase the service, the company agrees to
charge a service fee based on a percentage of the measured energy
savings."

As required by the M&V scenario, we assessed the performance of both the
conventional and the novel randomized M&V methods by estimating the
intervention energy savings for all valid buildings in the BG2 dataset.
By conducting such analysis, we aim to:

1.  Compare the energy saving estimation accuracy between the
    conventional and the randomized method. This study extends the
    comparison to a large sample of buildings, covering a variety of
    types and climate zones. The comparison metrics include both
    estimation accuracy and M&V completion timeline.

2.  Verify the enhanced robustness of the randomized method. By using
    realistic measurements from real-world buildings, which include
    various sources of noise, we aim to reflect the challenges faced by
    building analysts in real projects. As we will perform numercial
    simulation on the existing dataset, the ground-truth savings can be
    calculated as the reference for comparison.

3.  Open-source implementation of the proposed randomized M&V method
    using a public dataset. We ensured the reproducibility of the method
    by making the analysis code open source including randomized
    schedule generation, sequential statistical analysis, energy
    modeling and normalized saving calculation. In addition, we also
    included codes for extended use cases of the randomized method, such
    as changing sampling ratio and sampling intervals. Using the
    available open resources, building analysts should be able to
    seamlessly integrate and apply them in their own M&V projects.

@@

```{r flowchart, fig.cap = "Flow chat showing the methodology for comparing the estimated savings of randomized M&V with the conventional M&V", out.width="75%"}

@@

In this study, we extracted the electricity measurements from the BDG2
dataset and filter out all qualified buildings based on the following
criteria:

@@

3.  $\frac{abs(E_{2017} - E_{2016})}{E_{2016}}< 25\%$: any increase or
    decrease of building electricity usage in the second year should be
    less than 25% of that in the first year.

4.  Electric EUI \< 750 kWh/$m^2$: excludes buildings at the top 5%
    electrical energy usage intensity according to the statistics
    provided by the Building Performance Database
    [@lawrenceberkeleynationallab; @mathew_big-data_2015].

@@

```{r stablesite, fig.cap = "Site summary of the stable building subset (counts < 10 are omitted for visualization; left: aggragated counts of buildings for each type; right: breakdown building counts for each building type at each location)", fig.width=8, fig.height=6}

@@

ggarrange(p1, p2,
          ncol = 2, nrow = 1,
          widths = c(0.2, 1),
          common.legend = TRUE,
          legend = "bottom") +
  plot_annotation(title = "Summary of building numbers and types")
```

@@

1.  No statistical significant difference (P-value \> 0.05) between the
    two-year electricity usage, which are sometimes assumed to be the
    case for whole-building measurement and verification.

As a results, there are only 66 buildings were labeled in this 'stable'
subset (out of total of 600 buildings) implying that the assumption made
by conventional M&V is overly simplified. As mentioned in Section
\@ref(non-routine-events-impact), such variability is largely associated
with non-routine events and according to the statistics shown here, is
commonly observed in real applications.

@@

Figure \@ref(fig:chwst) shows the algorithm for the proposed control
intervention that reset the chiller supply temperature based on the
outdoor weather conditions, which can be commonly found in the
literature [@lee_chilled_2022; @congradac_recognition_2012]. For both
strategies, we assume that the chiller is activated when the outdoor
temperature exceeds 10°C. The baseline strategy, representing the
existing measurements from the dataset, operates with a constant water
supply temperature. The intervention strategy, as illustrated in the
figure, adjusts the water supply temperature dynamically, resetting it
from 7°C to 12°C.

```{r chwst, fig.cap = "Proposed intervention strategy: chilled water supply temperature reset based on outdoor temperature", fig.width=8, fig.height=4}

@@

We mapped the chilled water supply temperature reset to the electrical
energy savings as: $$
\text{savings} = mean(E_{CHW}) \times 25\% \times \text{perc_savings}
$$

@@

|                        |                                                                  |       |
|--------------|--------------------------------------------|--------------|
| Parameter              | Description                                                      | Value |
| $\beta_{temp}$         | \% savings from setting $T_{interv}$ 1 °C higher than $T_{base}$ | 0.08  |
| $\beta_{occ}$          | \% savings adjustment during occupied hours                      | 1.2   |
| $\beta_{\text{unocc}}$ | \% savings adjustment during unoccupied hours                    | 0.8   |

@@

We described more in detail the workflow of both conventional and
randomized M&V methods in the previous study [@raftery2024]. The
conventional method is a simple pre- and post- comparison on a 12-month
baseline and 12-month intervention timeline. The randomized method
defines sampling requirements as:

@@

-   A minimum and maximum of 12 and 108 weeks respectively. The
    randomized schedule covers the entire two-year period but stopping
    criteria enables an early stop at the end of satisfied blocking
    period.

@@

-   Test for no carryover effect using a t-test with a p-value not
    exceeding a defined significance threshold of 0.05.

@@

-   test with an equal sampling ratio (50% baseline, 50% intervention).

@@

The accuracy ($Acc$) is calculated as: $$
Acc = \frac{E_{baseline} - E_{intervention}}{E_{baseline}}-\frac{E'_{baseline} - E'_{intervention}}{E'_{baseline}}
$$where $E$ indicates true electricity usage and $E'$ indicates the
estimated electricity usage either through the conventional method or
the new randomized method. We present the distribution of savings
estimation accuracy using boxen plots, also known as letter plots, which
is an advanced variation of the box plot designed to extend beyond the
interquartile range (IQR) by progressively dividing the data into
smaller percentiles, revealing more detail in the tails of the
distribution. We set the division parameter to k = 4, meaning the entire
box area represents the data distribution from 6.25% to 93.75%. As the
steps move closer to the center line (50% median), the distribution
range progressively narrows with the next step represents 12.5% to
87.5%, followed by 25% to 75%. In addition, for each boxen plot, we also
showed the mean value of the distribution on top.

```{r abstract, fig.cap="Overall comparison results between the conventional M&V method and the proposed randomized M&V method (both at the stopping criteria and over a two-year period).", fig.height=12, fig.width=12}

@@

Additionally in the literature we found another M&V method study using
over 500+ commercial buildings, but their study scope was to compare a
variety of M&V baseline models from simple week mean to more complex
machine learning methods using the normalized mean bias error
(NMBE)[@granderson_accuracy_2016]. Since we also calculated the
normalized savings estimation error, we compared our normalized accuracy
of 600+ buildings with TOWT model and TMY weather with their TOWT model
prediction assessment in the following table. As each percentile is
closely matched, our results are verified.

|                 | Reference study results [@granderson_accuracy_2016] | Our study results      |
|-----------------|--------------------------------------|-----------------|
| 25th percentile | -5.85                                               | `r percentiles$Q1`     |
| 50th percentile | -1.25                                               | `r percentiles$Median` |
| 75th percentile | 3.86                                                | `r percentiles$Q3`     |

@@

```{r nosaving, fig.cap="Comparison between the two M&V methods in detecting no intervention effect", fig.height=6, fig.width=12}

@@

cvrmse <- df_model_acc_stable %>% 
  bind_rows(df_model_acc_variable) %>% 
  pivot_longer(c(conv, rand), 
               names_to = "method", 
               values_to = "cvrmse") %>% 
  mutate(method = as.factor(method), 
         method = recode_factor(method, 
                              "rand" = "Randomized", 
                              "conv" = "Conventional"))

>>>>

---
title: "Demonstrating the Reliability of Randomized Measurement and Verification for Switchable Control Retrofits Using a Large Open-Source Dataset"
author:
  - Aoyu Zou ^[Center for the Built Environment, University of California Berkeley, USA], Paul Raftery ^1^, Stefano Schiavon ^1^, Carlos Duarte ^1^, Gail Brager ^1^
abstract: Conventional measurement and verification (M&V) methods rely on pre- and post- retrofit comparison to estimate energy savings. They are often time-consuming and unreliable, especially when non-routine events including static changes and gradual changes denoted as operational drift occurs during the M&V process. Those events are unrelated to the intervention but significantly affect building energy consumption and thus when the analyst applies conventional M&V suggested by industry guidelines, the results can be largely confounded. In this study, we demonstrated that switchable interventions, such as most control retrofits, can benefit from a new M&V method which randomly samples whether to implement the baseline or the intervention strategy at a fixed interval (e.g., daily). We tested this novel randomized M&V method on a large public dataset covering various climate zones and commercial building types, using a virtual chilled water supply temperature reset based on outdoor weather as the intervention. Our results show that compared to the conventional method, the randomized method provides faster and more accurate savings estimations. Additionally we found that when non-routine events are present, the randomized M&V approach estimates savings that are much closer to the ground-truth savings than the conventional M&V method, demonstrating much improved reliability. We also assessed the impact of dropping samples affected by carryover when switching between strategies, and identify the optimal sampling interval for that scenario using this large dataset.

@@

# colors
ls_colors <- c("Baseline" = "grey10",
               "Intervention" = "#fdbb84",
               "Measured interv" = "#fdbb84",
               "Conventional" = "grey70", 
               "Randomized" = "#99d8c9", 
               "Randomized\n(24 months)" = "#6baed6",
               "Randomized\n(50/50)" = "#6baed6",
               "Randomized\n(20/80)" = "#6baed6", 
               "Randomized\n(10/90)" = "#6baed6", 
               "Daily\nsampling" = "#66c2a4",
               "2-day\nsampling" = "#66c2a4",
               "3-day\nsampling" = "#66c2a4",
               "7-day\nsampling" = "#66c2a4",
               "Dropped" = "#41ae76",
               "Kept" = "#006d2c",
               "Dropped (24 months)" = "#4292c6",
               "Kept (24 months)" = "#08519c",
               "Buildings finishing randomized M&V" =  "#99d8c9",
               "Buildings satisfying 80% TMY range" = "black", 
               "Buildings satisfying SPRT" = "black")

@@

Measurement and Verification (M&V) is the process of quantifying energy
savings from energy efficiency projects by comparing actual energy
consumption against a baseline, adjusting for factors like weather and
occupancy. This process ensures that improvements in energy performance
are accurately evaluated. In the United States, practitioners often
refer to ASHRAE Guideline 14, the International Performance Measurement
and Verification Protocol (IPMVP), and the Federal Energy Management
Program (FEMP) for standard guidelines [@doe_mv_2008;
@efficiency_valuation_organisation_international_2012;
@ashrae_ashrae_2023]. These guidelines outline standardized methods for
quantifying energy savings, whether through calibrated simulations or
monitored measurements for specific equipment or systems (isolation
methods) or for entire buildings (whole-building methods). In this
study, we will focus on the energy savings quantified at the
whole-building level where the measurements are obtained from utility
bills or whole-building meters, however the overall conclusions would
also apply to isolation methods. Typically, the process begins with
baseline measurements taken during the year before implementing the
energy-efficiency retrofit, followed by the same measurement procedure
during the year after the intervention. After collecting two years of
data, an M&V analyst fits an energy prediction model, using variables
such as outdoor temperature and time [@mathieu_quantifying_2011] to
project baseline energy consumption in the post-retrofit period. The
difference between the counter-factual baseline and the measured
intervention period energy consumption represents the energy savings.
One of the key drawbacks of this method is its reliance on a two-year
timeline to quantify savings, during which baseline measurements can
become outdated due to changes in building performance caused by
non-routine events unrelated to the intervention. Specifically, an M&V
analyst should account for two types of baseline changes: (1) static
changes, which are known building operational changes: such as
renovation, equipment addition or removal; and (2) gradual changes,
which often occur subtly and unknown, such as the incremental adoption
of LED lighting or more efficient plug loads. Identifying and
quantifying these changes are particularly labor-intensive and
challenging. Consequently, those limitations reduce the feasibility of
M&V and complicates the quantification of estimation uncertainty, thus
impacting the accuracy and timeliness of savings assessments.

@@

To address the limitations of conventional M&V methods and the
challenges posed by non-routine events, we proposed a M&V method that
adopts the randomized crossover design [@raftery2024], a gold standard
from medical, agricultural and online controlled experimental (i.e. A/B
testing) studies [@blackston2019; @duan2013; @gupta2019; @mie2022;
@munro; @raseduzzaman2017]. An early application of this randomized
design to the field of M&V in buildings was first used in
[@raftery2018], which compared the energy performance on two different
supply air temperature reset strategies in a building. Subsequent
improvements to the M&V method include adding a sequential evaluation
framework and defined stopping criteria to end the M&V period early
after reaching a target level of uncertainty, among other improvements,
such as using blocked randomization by weekday. The full framework is
detailed in a previously published study [@raftery2024] along with a
case study application to a real building, and other example use cases.
In summary, this method provides M&V analysts with a randomized schedule
that alternates between baseline and intervention implementation while
ensuring balanced sampling across days of the week and seasons. For
example, given a 10-week M&V period for 1 intervention, the balanced
randomized schedule would equally sample 5 Mondays with the baseline
strategy in operation and 5 Mondays with the intervention strategy in
operation. The limitation of this method is that it is only applicable
to a subset of retrofit projects where interventions can be easily
switched on and off. However, for all applicable use cases, it allows
analysts to detect energy savings sequentially shortly after the test
begins. In addition, once the desired estimate uncertainty target is
achieved, analysts can terminate the M&V study and switch to 100%
intervention. The key advantage of randomization is that if control
strategies are sampled with equal probability, the influence of other
confounding factors such as occupancy change (commonly known as one of
the non-routine events) and other more subtle long term changes in
building energy use are likely to be evenly distributed among
measurements, leading to a more accurate and unbiased assessment of the
intervention effect.

@@

To demonstrate and differences between the two M&V methods, the Building
Genome Dataset 2 (BGD2) is used, which is an extensive open-access
dataset designed to advance research and development in building energy
efficiency and control strategies acting as a test-bed for modeling,
simulation, and algorithm development [@miller_building_2020]. BGD2
contains over 1000 buildings' metadata and realistic operational
measurements between 2016 and 2017 from across North America and Europe,
making it one of the most comprehensive collections of building-related
data available for scientific use. The dataset includes various
commercial building types such as offices, education facilities, public,
and retail buildings, and provides detailed information on their
physical characteristics (e.g. energy ratings, heating types and floor
area) and hourly measurements of chilled and hot water, electricity, gas
usage as well as site outdoor weather conditions. In this study, we
mostly use the metered electricity and outdoor weather measurements for
running the M&V methods.

@@

ASHRAE Guideline 14 and the IPMVP provide options for whole-building M&V
use cases. For code compliance using the prescriptive option, 12 months
of baseline and 12 months of post-retrofit measurements are required. In
addition, the expected savings should be larger than 10% and the
baseline model fitting accuracy should have CV(RMSE) lower than 20%.
Thus, most research related to M&V for whole-building approach focuses
on the accuracy of baseline modeling, exploring model performance from
simple regression models to more complex machine learning techniques.
One study reviewed various models suitable for M&V applications as well
as selected input features [@alrobaie_review_2022] and another study
provided a definitive methodology to apply machine learning models for
M&V use cases [@gallagher_development_2018]. In addition, a few studies
investigated the critical performance metrics to evaluate the developed
baseline models [@granderson_automated_2015;
@granderson_development_2014] and compared a variety of models using
those metrics [@granderson_accuracy_2016]. These studies made
significant contributions by emphasizing the uncertainty associated with
the model-fitting process, a key factor in accurately determining energy
savings. Furthermore, other researchers addressed this issue by
leveraging statistical formulation and inference to improve baseline
energy models [@burkhart_measurement_2014; @heo_calibration_2012;
@walter_uncertainty_2014]. However, a gap still remains in the
literature regarding the rigorous quantification of uncertainties
directly associated with calculated savings, for instance, accounting
for the potential bias that the baseline model might deteriorate (i.e.
become 'stale') over the extended period of pre- and post-analysis.

@@

Non-routine events in M&V commonly refer to unexpected changes in a
building that influence a building's energy usage and are typically
unrelated to the intervention strategy. Current guidelines define
non-routine events as 'static factors' that needed to be adjusted after
projecting baseline in the post-retrofit period [@ashrae_ashrae_2023;
@efficiency_valuation_organisation_international_2012]. A common
non-routine event in energy-saving M&V projects is a change in occupancy
or a significant shift in occupant behavior, equipment run time, and
operating conditions (e.g. set points, lighting, and ventilation
levels). However, current standards or guidelines only provide a general
approach for consideration. For example, IPMVP requires the facility
owner and the M&V analyst to periodically perform inspections of all
equipment and operations during the reporting period, which is
labor-intensive and error-prone. ASHRAE guideline 14 recommends
performing engineering calculations or computer software simulations to
adjust the post-retrofit baseline. Additionally, it is relatively rare
to have access to all measurements needed for adjustments such as
occupancy, and thus the analyst normally assumes those factors remain
unchanged throughout the study. For example, some studies uses linear
interpolation to estimate counter-factual baseline for demand response
program [@beil_round-trip_2015; @keskar_commercial_2020]. Consequently,
one study points out that it is inaccurate to assume no change in the
operating conditions during the response period
[@huang_experimental_2023]. Other studies realized such limitation in
field study, and emphasized the importance of requiring matched groups
to control for exogenous factors beyond weather differences when
comparing between baseline and intervention
[@demand_side_analytics_population_2022; @huang_experimental_2023].

Overall, those non-routine events mostly refer to clearly observable
changes, but there are also more gradual and subtle changes in
buildings, which are hardly noticed. We define those gradual changes as
'operational drift' in this study. One likely encountered operational
drift situation is the filter clogging in air handling units due to
particle accumulation. This causes supply fans to gradually consume more
energy to maintain the required duct static pressure [@feng_newly_2019;
@zhai_full-scale_2017]. If M&V analysts are unaware of such changes and
lack an appropriate adjustment method (e.g., replacing filters before
the intervention begins), the savings could be overestimated as the
decreased energy use is incorrectly attributed to the intervention
rather than the filter replacement. An operational drift has not been
formally defined in the literature or standards, but in this study, we
consider it as a special type of non-routine events as it also
represents a change in facility operation but in a longer term.

@@

As mentioned, the goal of an M&V project is to determine the effect,
typically energy savings, of some intervention in the building. In this
study, we focus on switchable interventions and an example of such an
intervention is a control strategy that adjusts the chilled water
plant's supply water temperature, which can be commonly found in the
literature [@duarte_field_2023; @lee_simulationoptimization_2012;
@qiu_chilled_2022; @jin2007; @taylor_optimizing_2012]. In our case, we
make it even simpler by adjusting the setpoint based on outdoor weather
conditions with more description shown in Section
\@ref(apply-virtue-intervention). By conducting such an analysis, we aim
to:

1.  Determine, using a large sample of buildings, how much more
    accurately the randomized M&V method would estimate the savings of
    the proposed intervention compared to the conventional M&V, and how
    much more quickly it would reach a result.

2.  Determine how much more robust the randomized method is compared to
    the conventional method when static non-routine events and
    operational drift are present in the dataset. The BDG2 dataset
    contains real-world building energy usage data collected over two
    years, inherently contains varying degrees of usage changes. By
    virtually introducing intervention effects into the existing
    dataset, we can calculate the ground-truth energy savings, which
    serve as a reference point for method comparison.

3.  Open-source implementation of the proposed randomized M&V method. We
    ensured the reproducibility of the method by making the analysis
    code open source including randomized schedule generation,
    sequential statistical analysis, energy modeling and normalized
    saving calculation. In addition, we also included code for extended
    use cases of the randomized method, such as changing sampling ratio
    and sampling intervals. Using the available open resources, building
    analysts should be able to seamlessly integrate and apply them in
    their own M&V projects.

@@

```{r flowchart, fig.cap = "Flow chart showing the methodology for comparing the estimated savings of randomized M&V with the conventional M&V", out.width="75%"}

@@

We extracted the electricity measurements from the BDG2 dataset and
filter out all qualified buildings based on the following criteria:

@@

3.  $\frac{abs(E_{2017} - E_{2016})}{E_{2016}}< 25\%$: any increase or
    decrease in building electricity usage in the second year should be
    less than 25% of that in the first year.

4.  Electric EUI \< 750 kWh/$m^2$: excludes buildings at the top 5% of
    electrical energy usage intensity according to the statistics
    provided by the Building Performance Database
    [@lawrenceberkeleynationallab; @mathew_big-data_2015].

@@

```{r stablesite, fig.cap = "Site summary of the filtered buildings from BDG2 dataset (counts < 10 are omitted for visualization; left: aggregated counts of buildings for each type; right: breakdown building counts for each building type at each location)", fig.width=8, fig.height=6}

@@

ggarrange(p1, p2,
          ncol = 2, nrow = 1,
          widths = c(0.2, 1),
          common.legend = TRUE,
          legend = "bottom") +
  plot_annotation(title = "Summary of filtered buildings and types")
```

@@

1.  No statistically significant difference (P-value \> 0.05) between
    the two-year electricity usage, which are sometimes assumed to be
    the case for whole-building measurement and verification.

As a result, there are only 66 buildings were labeled in this 'stable'
subset (out of total of 600 buildings) implying that the assumption that
buildings are typically stable over the timeframe involved in
conventional M&V is rarely true. As mentioned in Section
\@ref(non-routine-events-impact), such variability is largely associated
with non-routine events, especially operational drifts since the
assessment last for two years. And according to the statistics shown
here, it is a typical case observed in real buildings applications.

@@

Figure \@ref(fig:chwst) shows the algorithm for the proposed control
intervention that reset the chiller supply temperature based on the
outdoor weather conditions, which can be commonly found in the
literature [@lee_chilled_2022; @congradac_recognition_2012]. For both
strategies, we assume that the chiller operates when the outdoor
temperature exceeds 10°C. The baseline strategy, representing the
existing measurements from the dataset, operates with a constant water
supply temperature. The intervention strategy, as illustrated in the
figure, adjusts the water supply temperature dynamically, resetting it
from 7°C to 12°C.

```{r chwst, fig.cap = "Proposed intervention strategy: chilled water supply temperature reset based on outdoor temperature.", fig.width=8, fig.height=4}

@@

We used a simple relationship to map the chilled water supply
temperature reset to the electrical energy savings: $$
\text{savings} = mean(E_{CHW}) \times 25\% \times \text{perc_savings}
$$

@@

|                        |                                                                  |       |
|------------------|------------------------------------|------------------|
| Parameter              | Description                                                      | Value |
| $\beta_{temp}$         | \% savings from setting $T_{interv}$ 1 °C higher than $T_{base}$ | 0.08  |
| $\beta_{occ}$          | \% savings adjustment during occupied hours                      | 1.2   |
| $\beta_{\text{unocc}}$ | \% savings adjustment during unoccupied hours                    | 0.8   |

@@

We described in detail the workflow of both conventional and randomized
M&V methods in the previous study [@raftery2024]. The conventional
method is a pre- and post- comparison on a 12-month baseline and
12-month intervention timeline. The randomized method defines sampling
requirements as:

@@

-   A minimum and maximum of 12 and 108 weeks respectively. The
    randomized schedule covers the entire two-year period but stopping
    criteria enables an early stop at the end of the satisfied blocking
    period.

@@

-   No carryover effect using a t-test with a p-value not exceeding a
    defined significance threshold of 0.05.

@@

-   Sample strategies at an equal sampling ratio (50% baseline, 50%
    intervention).

@@

The accuracy ($Acc$) is calculated as: $$
Acc = \frac{E_{baseline} - E_{intervention}}{E_{baseline}}-\frac{E'_{baseline} - E'_{intervention}}{E'_{baseline}}
$$where $E$ indicates true electricity usage and $E'$ indicates the
estimated electricity usage either through the conventional method or
the new randomized method. We present the distribution of savings
estimation accuracy using boxen plots, also known as letter plots, which
is an advanced variation of the box plot designed to extend beyond the
interquartile range (IQR) by progressively dividing the data into
smaller percentiles, revealing more detail in the tails of the
distribution. We set the division parameter to k = 4, meaning the entire
box area represents the data distribution from 6.25% to 93.75%. As the
steps move closer to the center line (50% median), the distribution
range progressively narrows with the next step representing 12.5% to
87.5%, followed by 25% to 75%. In addition, for each boxen plot, we also
showed the mean value of the distribution on top for reference.

```{r abstract, fig.cap="Overall comparison results between the conventional M&V method and the proposed randomized M&V method (both upon reaching the stopping criteria for each case, and over a fixed two-year period).", fig.height=12, fig.width=12}

@@

Additionally in the literature, we found another M&V study using over
500+ commercial buildings, but their study scope was to compare the
predictive performance of a variety of M&V baseline models from simple
weekly mean to more complex machine learning methods using the
normalized mean bias error (NMBE) [@granderson_accuracy_2016]. Since we
also calculated the normalized savings estimation error, we compared our
normalized accuracy for 600+ buildings with the TOWT model and TMY
weather with their TOWT model prediction assessment in the following
table, highlighting that the model and overall building datasets are
performing similarly.

|                 | Reference study results [@granderson_accuracy_2016] | Our study results      |
|-------------------|-----------------------------------|-------------------|
| 25th percentile | -5.85                                               | `r percentiles$Q1`     |
| 50th percentile | -1.25                                               | `r percentiles$Median` |
| 75th percentile | 3.86                                                | `r percentiles$Q3`     |

@@

```{r nosaving, fig.cap="Comparison between the two M&V methods in detecting no intervention effect when buildings are subject to static non-rountine events and gradual operational drift.", fig.height=6, fig.width=12}

@@

cvrmse <- df_model_acc_stable %>% 
  bind_rows(df_model_acc_variable) %>% 
  pivot_longer(c(conv, rand), 
               names_to = "method", 
               values_to = "cvrmse") %>% 
  mutate(method = as.factor(method), 
         method = recode_factor(method, 
                              "rand" = "Randomized\n(24 months)", 
                              "conv" = "Conventional"))

result <- t.test(filter(cvrmse, method == "Randomized\n(24 months)")$cvrmse, 
                 filter(cvrmse, method == "Conventional")$cvrmse, 
                 exact = TRUE,
                 alternative = "less",
                 mu = 0,
                 conf.int = TRUE, 
                 conf.level = 0.90,
                 var.equal = TRUE)$p.value
